---
title: "FinalProjectWeek4"
author: "Mateo Córdoba Toro"
date: "25/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message=FALSE, results='hide'}
library(dplyr)
library(caret)


training <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv")
test <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv")

training <- training[,!(names(training) %in% c('X','user_name',"raw_timestamp_part_1","raw_timestamp_part_2") )]
test <- test[,!(names(training) %in% c('X','user_name',"raw_timestamp_part_1","raw_timestamp_part_2") )]
```


The first part of the process is to perform a data cleaning, a function is built to detect the percentage of missing values in the data set.
```{r}

na.percent <- function(x){
  sum(is.na(x))/length(x)
}


x <- sapply(training, na.percent)
x <- names(x[x>0])
x
```

It is found that many of the variables  present an exaggerated amount of missing values, approximately 97%, which is why they are omitted. Additionally, variables that provide very little variance are filtered.

```{r}
training <- training[,!(names(training)%in%x)]
testing <- test[,!(names(training)%in%x)]
vars <- names(training)[nearZeroVar(training)]
vars
```

After performing the cleaning process, the data within the data set is divided into a training set and a validation set, since the amount of data that is presented for training is very few.


```{r}
training <- training[,!(names(training) %in% vars)]
testing  <- test[,!(names(training) %in% vars)]
set.seed(123)

index <- createDataPartition(training$classe, p = 0.5, list = FALSE)
sub.train <- training[index,]
sub.test  <- training[-index,]
```

# MODEL1: DESICION TREE

The first model to be estimated is a decision tree model, The decision matrix shows the degree of successes it presents, showing a high precision in the classification.

```{r}
model1 <- rpart::rpart(classe~., data = sub.train, method = "class")
pred1  <- predict(model1, sub.test, type = "class")
table(pred1,sub.test$classe)
```
As can be seen, the model is almost correct in its entirety, with a accurancy of 83%.

```{r}
sum(diag(table(pred1,sub.test$classe)))/length(sub.test$classe)

```



# MODEL2: RANDON FOREST 

The second model to estimate a random forest, with cross validation.

```{r}
controlRF <- trainControl(method="cv", number=5)
model2 <- train(classe ~ ., data=sub.train, method="rf", trControl = controlRF)
```


It can be seen that the precision of the model improves substantially, placing it at 99% precision.

```{r}
rfPrediction <- predict(model2, sub.test)
table(sub.test$classe, rfPrediction)
sum(diag(table(rfPrediction,sub.test$classe)))/length(sub.test$classe)

```


```{r}
dt.test <- predict(model1,testing, type = "class")
rfPrediction.test <- predict(model2, testing)

data.frame(DECISION.TREE = as.character(dt.test),
           RANDOM.FOREST = as.character(rfPrediction.test),
           COINCIDENCE = dt.test == rfPrediction.test)
```

#Conclusion.

The model estimated by random forest is taken since in the validation it presents a higher precision than model 1, the models represent a 75% coincidence, so the results of the decision tree also seem to be good but not as precise as the estimated by random forest.




